---
output:
  github_document:
    toc: TRUE
---

```{r setup, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  dpi = 300,
  out.width = "100%",
  comment = "#>",
  fig.path = "man/figures/README-"
)
## so jittered figs don't always appear to be changed
set.seed(1)
```

## gapminderplus

This package extends the `gapminder` package by Jennifer Bryan with one additional data frame `country_geo`, which contains latitude, longitude, and surface area information for the countries in the `gapminder` data set. Information about the original `gapminder` package can be found [here](https://github.com/jennybc/gapminder).

Instead of using vignettes for illustration, I will show here how the new `country_geo` data frame can be used with the `gapminder` data set.

### Install and test drive

Install `gapminderplus` from GitHub:

```{r eval = FALSE}
devtools::install_github("STAT545-UBC-students/hw07-YuHsiangLo/gapminder")
```

Load it and test drive with some data aggregation and plotting:

```{r test-drive, message = FALSE, warning = FALSE}
library(gapminderplus)
library(tidyverse)

ggplot(gapminder, aes(x = continent, y = lifeExp, color = continent)) +
  geom_violin() +
  geom_jitter(alpha = 0.25)
```

### Latitudes, longitudes, and surface areas

The `country_geo` data frame provides the average latitudes and longitudes and surface areas for all the countries in the `gapminder` data frame. This additional information can be used to generate more informative plots, for example, with the `ggmap` package.

```{r, message=FALSE, warning=FALSE}
library(ggmap)

gapminder %>%
  filter(year == 2007) %>%
  left_join(., country_geo, by = "country") %>%
  mutate(density = pop / area) %>%
  ggplot(aes(x = longitude, y = latitude, size = density, color = lifeExp)) +
  theme_bw() +
  borders("world", colour = "gray80", fill = "gray75") +
  geom_point() +
  scale_size_area() +
  scale_size_continuous(name = "Pop. density") +
  scale_color_continuous(name = "Life exp.") +
  labs(title = "Population density and life expectancy in 2007") +
  theme(plot.title = element_text(hjust = 0.5),
        axis.title = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank())
```

### More interesting use of the `country_geo` data frame

- We can inspect the relationship between latitude and GDP per capita over the years. Using the `gganimate` package, we can even animate the changes.

```{r, message=FALSE, warning=FALSE}
library(MASS)  # for robust regression
library(gganimate)

gapminder %>%
  left_join(., country_geo, by = "country") %>%
  ggplot(aes(x = abs(latitude), y = log(gdpPercap))) +
  theme_bw() +
  geom_point(aes(color = continent)) +
  geom_smooth(method = rlm, se = FALSE) +
  labs(title = 'Year: {frame_time}', x = 'Absolute latitude', y = 'GDP per capita') +
  theme(plot.title = element_text(hjust = 0.5),
        legend.title = element_blank()) +
  transition_time(year) +
  ease_aes('linear')
```

We can see that there is a positive relation between latitude and GDP per capita, with countries in higher latitudes also having higher GDP per capita. Also, this trend seemed to grow more prominent over the years.

- Now instead of the concept of GDP per capita, let us define another measure called *GPD per square kilometer*, which quantifies how much money was generated per square kilometer by each country. We can then find the top five countries that have the highest GDP per square kilometer in each year.

```{r}
gapminder %>%
  left_join(., country_geo, by = "country") %>%
  mutate(gdp = gdpPercap * pop,
         gdpPerSqKm = gdp/area) %>%
  group_by(year) %>%
  top_n(5, gdpPerSqKm) %>%
  ungroup() %>%
  arrange(year, gdpPerSqKm) %>%
  mutate(order = row_number()) %>%
  ggplot(aes(x = order, y = gdpPerSqKm, group = year, fill = country)) +
  theme_bw() +
  geom_bar(stat = 'identity') +
  scale_y_log10() +
  geom_text(aes(label = country, y = 1), hjust = 0, size = 3) +
  facet_wrap(~ year, scales = 'free_y') +
  labs(x = 'Country', y = 'GDP per square kilometer') +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position = 'none') +
  coord_flip()
```

This plot shows us that **bigger isn't always better**: small countries can also generate a lot of money. It should be noted, though, that these countries tend to be densely populated as well. So perhaps it is not that surprising after all --- more people on a small land, more money on a small land!